COVID19 - Forecast and predictions using a time dependent SEIR model - Veneto Region

### 2020-03-21 BYM_model/INLA_Veneto_html.Rmd
Add home page, navigation and theme

```{r}
####specification
formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
  f(t,model="rw2",constr = FALSE) +
  f(t2,model="iid") +
  f(ID2,model="besag", graph=file.adj,
    group=t3,
    control.group=list(model="rw2"))
#estimation
fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_st4)
```

>>>>

```{r}
####specification
formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
  f(t,model="rw2") +
  f(t2,model="iid") +
  f(ID2,model="besag", graph=file.adj,
    group=t3,
    control.group=list(model="rw2"))
#estimation
fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_st4)
```

### 2020-03-21 draft_analysis_Italy_html.Rmd
Add home page, navigation and theme

## Load dataset
```{r}

rm(list=ls())
###import italian updated dataset 
  dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
days<-dim(dat_csv)[1]
dat_csv$t<-1:days
# The total number of epidemic day is
days

```
Several outcomes can be potentially monitored, that is

@@

With the aim of predicting the future number of COVID19 cases on the
basis of the actual data, we used a SEIR model applied to the COVID19
epidemic in Italy.

@@

The model divides a population of hosts into three classes:
susceptible, infected, recovered.
The model describes how the portion of the population in each of these classes changes with time.
Births are modeled as flows from "elsewhere" into the susceptible class;
deaths are modeled as flows from the $S$, $I$, or $R$ compartment into "elsewhere".
If $S$, $I$, and $R$ refer to the numbers of individuals in each compartment, then these **state variables** change according to the following system of differential equations:
$$\begin{aligned}
\frac{d}{dt}S(t) &= B(t)-\lambda\,S(t)-\mu\,S(t)\\
\frac{d}{dt}I(t) &= \lambda\,S(t)-\gamma\,I(t)-\mu\,I(t)\\
\frac{d}{dt}R(t) &= \gamma\,I(t)-\mu\,R(t).\\
\end{aligned}$$
Here, $B$ is the crude birth rate (births per unit time), $\mu$ is the death rate and $\gamma$ is the recovery rate.
We'll assume that the force of infection, $\lambda$, for a constant population $N$
$$\lambda = \beta\,\frac{I}{N},$$
so that the risk of infection a susceptible faces is proportional to the *prevalence* (the fraction of the population that is infected).
This is known as the assumption of frequency-dependent transmission.

@@

$$
\log(I(t))= \beta_0 + \beta_1  t +e_t
$$

@@

The incubation period $1/ \gamma$ for the coronavirus is in mean 5.1 days with a range from 2-14 days. Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/. However, the incubation period is used for epidemic diseases that causes the immediate home isolation of infected subjects.

In the calculation we considered an "incubation period" of 14 days for two reasons:  
1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14. A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 
2) 14 days is the worst scenario because in this period low impact COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).

@@

```{r R0 trend}
#calculate r0 based with a mobile window of 5 days
#vector for beta and standard deviation
beta_vec<-NULL
sd_vec<-NULL
#for cycle for R0 estimates from days-2 to days+2
for (i in 3:(days-2)){
fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(i-2):(i+2),])
beta_vec<-c(beta_vec,coef(fit)[2])
sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
}

@@

mean  <- 1+(beta_vec*14)
lower <- 1+((beta_vec-1.96*sd_vec)*14)
upper <- 1+((beta_vec+1.96*sd_vec)*14)

@@

time<-3:(days-2)
beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=gaussian)
forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
beta.predict<-predict 
r0.predict<-beta.predict
r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*14+1
# plot the points (actual observations), regression line, and confidence interval
p <- ggplot(r0.predict, aes(date,beta_vec))
p <- p + geom_point() +labs(x="Date",y="R0 value") 
p <- p + geom_line(aes(date,fit))
p <- p + geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)
p
```

R0 passes from a value of `r round(r0.predict[1,1],2)`   in the initial phase to an  estimated value of `r round(r0.predict[days+forecast,3],2)` at the ending of the 14-days forecast.

We want to make a short term forecast (14 days) with 3 scenario, based on the number of exposed people:

-Scenario 1: 10 exposed people for each COVID-19 case (no home restrictions made or even no effects)

-Scenario 2: 5 exposed people for each COVID-19 case (-50% exposed people)

-Scenario 3: 3 exposed people for each COVID-19 case  (-70% exposed people)

We made a forecast by means of a SEIR model fixing a series of initial status:  
- S0=N, the size of Italian population   
- E= f * I_start (with f a fixed factor of the previous scenario)  
- I_start: initial number of COVID-19 cases  
- R_start: initial number of recovered

and parameters:  
- beta: gamma+slope $\beta_1$
- gamma= 1/duration (rate of infection duration of COVID-19, 14 days)  
- sigma0: the coronavirus transmission rate (half of flu epidemic)   
- mu0: the overall mortality rate

```{r initation plot, echo=TRUE}
# initial number of infectus
I_start<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I_start
# initial number of recovered
R_start<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R_start
# italian poulation
N=60480000
# duration of COVID19 
duration<-14
#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
sigma0<-0.05
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```
We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:

<img src="http://www.public.asu.edu/~hnesse/classes/seireqn.png"/>  
<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>  
Respect to the previous formulation, the compartment of exposed people (E) was inserted between the susceptible and infected compartments.  
(https://en.wikipedia.org/wiki/Bayesian_structural_time_series)

We perfom simulation on the trend of beta coefficient by means of a Bayesian Structural Time Series using the library bsts of R.    
We estimate a BSTS model specifing a local linear trend ans 1000 simulation.

@@

```{r bsts model fit and prediction, echo=TRUE}
# Bayesian Structural Time Series
set.seed(123)
ss <- AddLocalLinearTrend(list(), beta_vec)
model1 <- bsts(beta_vec,
               state.specification = ss,
               niter = 1000)
par(mfrow = c(1,1))
plot(model1, "components", ylab="Beta coefficient", xlab="days")
#previsioni
pred1 <- predict(model1, horizon = 16, burn = 100)
par(mfrow = c(1,1))
plot(pred1 , ylab="Beta coefficient",main="Data and predictions")
# matrix of beta coefficients
coef<-(pred1$distribution[,3:16])
par(mfrow = c(1,1))
```
For each vectors of simulated beta coefficients we perform a SEIR model (one for each scenario).  
We save the results and plot a credible interval at 90% (5-95% percentile).

```{r  scenario plot , echo=TRUE}

seir1_sim<-seir2_sim<-seir3_sim<-NULL
for(s in 1:dim(coef)[1]){
  # average number of single connections of an infected person
  # less contacts, less probability of new infections
  # we keep constant the other parameters
  forecast<-14
  seir1<-seir2<-seir3<-NULL
  for(i in 1:forecast){
    parameters <- c(mu = mu0, beta = matrix(coef[s,i])+1/duration, sigma = sigma0, gamma = 1/duration)
    f1<-10
    if( i==1) initials <- c(S = 0.95, E = (f1*I_start/N), I = I_start/N, R = R_start/N)
    if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
    seir1_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
    seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
    f2<-5
    if( i==1) initials <- c(S = 0.95, E = (f2*I_start/N), I = I_start/N, R = R_start/N)
    if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
    seir2_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
    seir2 <- rbind(seir2,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
    f3<-3
    if( i==1) initials <- c(S = 0.95, E = (f3*I_start/N), I = I_start/N, R = R_start/N)
    if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])
    seir3_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
    seir3 <- rbind(seir3,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
  }
  seir1_sim<-rbind(seir1_sim,cbind(rep(s,forecast),seir1))
  seir2_sim<-rbind(seir2_sim,cbind(rep(s,forecast),seir2))
  seir3_sim<-rbind(seir3_sim,cbind(rep(s,forecast),seir3))
}
seir1_sim[,2]<-seir2_sim[,2]<-seir3_sim[,2]<-rep(1:forecast,dim(coef)[1])
colnames(seir1_sim)[1]<-colnames(seir2_sim)[1]<-colnames(seir3_sim)[1]<-"sim"

### confidence limits
I_seir_med<-tapply(seir1_sim$I,seir1_sim$time,median)
limits<-c(0.05,0.95)
confidence<-diff(limits)
I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=limits[1])
I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=limits[2])

days.before<-date[1:days]
days.ahead<-date[(days+1):(days+forecast)]
step.ahead<-forecast+1
mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Italy: Scenario 1 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p

### confidence limits - scenario 2
I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=limits[1])
I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=limits[2])

mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Italy: Scenario 2 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p

### confidence limits - scenario 3
I_seir_med<-tapply(seir3_sim$I,seir3_sim$time,median)
I_seir_lwr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=limits[1])
I_seir_upr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=limits[2])

mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Italy: Scenario 3 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```

The 3 scenarios show different numbers. 
If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N)[forecast]`).

In the next plot the cumulative number of infected.  
At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N+tapply(seir2_sim$R,seir2_sim$time,median)*N)[forecast]`).

>>>>

## Load dataset
```{r}
rm(list=ls())
###import italian updated dataset 
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
days<-dim(dat_csv)[1]
dat_csv$t<-1:days
# The total number of epidemic day is
days
```
Several outcomes can be potentially monitored, that is

@@

With the aim of predicting the future number of COVID19 cases on the
basis of the actual data, we used a SEIR model applied to the COVID19
epidemic to Italy

@@

The model divides a population of hosts into three classes:
susceptible, infected, recovered.
The model describes how the portion of the population in each of these classes changes with time.
Births are modeled as flows from "elsewhere" into the susceptible class;
deaths are modeled as flows from the $S$, $I$, or $R$ compartment into "elsewhere".
If $S$, $I$, and $R$ refer to the numbers of individuals in each compartment, then these **state variables** change according to the following system of differential equations:
$$\begin{aligned}
\frac{d}{dt}S(t) &= B(t)-\lambda\,S(t)-\mu\,S(t)\\
\frac{d}{dt}I(t) &= \lambda\,S(t)-\gamma\,I(t)-\mu\,I(t)\\
\frac{d}{dt}R(t) &= \gamma\,I(t)-\mu\,R(t).\\
\end{aligned}$$
Here, $B$ is the crude birth rate (births per unit time), $\mu$ is the death rate and $\gamma$ is the recovery rate.
We'll assume that the force of infection, $\beta$, for a constant population $N$
$$\lambda = \beta\,\frac{I}{N},$$
so that the risk of infection a susceptible faces is proportional to the *prevalence* (the fraction of the population that is infected).
This is known as the assumption of frequency-dependent transmission.

@@

$$
\log(I(t))= \beta_0 + \beta_1  t +e_t.
$$

@@

The incubation period  for the coronavirus is in mean 5.1 days with a range from 2-14 days. Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/. However, the incubation period is used for epidemic diseases that causes the immediate home isolation of infected subjects. The duration of the diseases is about 2 weeks.

However, in the calculation of R0 we considered an infectious period of 21 days for two reasons:  
1) the majority of cases is asymptomatic;
2) 21 days is the worst scenario because in this period COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).

@@

```{r R0 trend}
#calculate r0 based with a mobile window of 5 days
#vector for beta and standard deviation
duration<-21
beta_vec<-NULL
sd_vec<-NULL
#for cycle for R0 estimates from days-2 to days+2
for (i in 3:(days-2)){
fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(i-2):(i+2),])
beta_vec<-c(beta_vec,coef(fit)[2])
sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
}

@@

mean  <- 1+(beta_vec*duration)
lower <- 1+((beta_vec-1.96*sd_vec)*duration)
upper <- 1+((beta_vec+1.96*sd_vec)*duration)

@@

time<-3:(days-2)
beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=gaussian)
forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
beta.predict<-predict 
r0.predict<-beta.predict
r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*21+1
# plot the points (actual observations), regression line, and confidence interval
p <- ggplot(r0.predict, aes(date,beta_vec))
p <- p + geom_point() +labs(x="Date",y="R0 value") 
p <- p + geom_line(aes(date,fit))
p <- p + geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)
p
```

R0 passes from a value of `r round(r0.predict[1,1],2)`   in the initial phase to an  estimated value of `r round(r0.predict[days+forecast,3],2)` at the ending of the 14-days forecast.  
We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:

$$\begin{aligned}
\frac{d}{dt}S(t) &= \mu (N-S)-\beta \frac{SI}{N}-\nu S\\  
\frac{d}{dt}E(t) &= \beta \frac{SI}{N}-(\mu+\sigma) E    \\
\frac{d}{dt}I(t) &=\sigma E- (\mu+\gamma)I \\
\frac{d}{dt}R(t) &= \gamma\,I(t)-\mu\,R(t)+\nu S.\\
\end{aligned}$$

<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>

Respect to the previous formulation, the compartment of exposed people (E) was inserted between the susceptible and infected compartments.  
(https://en.wikipedia.org/wiki/Bayesian_structural_time_series).
The parameter $\nu$ is the rate of vaccination.

We want to make a short term forecast (14 days).

We made a forecast by means of a SEIR model fixing a series of initial status:
*S:N, the size of Italian population  
*E: The number of exposed people (unknown)  
*I_start: initial number of COVID-19 cases     
*R_start: initial number of recovered

and parameters:  
- beta: gamma + slope
- gamma= 1/duration of diseases (duration=21 days)
- sigma: the coronavirus transmission rate 
- mu0: the overall mortality rate

```{r ini parameters plot, echo=TRUE}
# initial number of infectus
I_start<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I_start
# initial number of recovered, based on the proportion of discharged from the health services
prop<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]/dat_csv$totale_ospedalizzati[dim(dat_csv)[1]]
R_start<-prop*dat_csv$totale_casi[dim(dat_csv)[1]]; R_start
# Italian population
N=60480000
# duration of COVID19 as the sum of incubation (5-14 days) and the duration of diseases (about 20 days)
duration<-21
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```

We try to estimates sigma (from exposed to infected) and the number of exposed population on the basis of the last 5 days of observed number of total infected people.

We suppose that the number of total infected $I(t)$ can be the realization of a Poisson random variable
$$
I(t) \sim Poisson (\theta_t)
$$
where $\theta_t$ is the unknown mean of the random process at the time $t$.

The Poisson distribution has a fixed variance that is equal to the mean; in the next part we try to estimates the parameters considering a Negative Binomial distribution which allows for overdispersion.

However, the parameter $\theta_t$ can be supposed to be a realization of a SEIR model with initial parameters $\theta=(\mu,\beta,\sigma, \gamma)$  and initial status $S=(S_{start},E_{start},I_{start},R_{start})$.

The parameter $\sigma$ and the fraction of exposed people ($E_{start}$) are our parameters of interest; we estimate $\sigma$ and $E_{start}$ minimizing the log likelihood of the described Poisson distribution (or Negative Binomial one).

```{r estimation of gamma and exposed, echo=TRUE}
# number of the last considered days for calibration
last<-5
#I_start
I_1<-dat_csv$totale_attualmente_positivi[(days-last)]
#R_start
R_1<-prop*dat_csv$totale_casi[(days-last)]
#loglikelihood function
LogLikfun <- function (initials,parameters,obs) {
n<-length(obs)
N=60480000
seir_fit <- SEIR(pars = parameters, init = initials, time = 0:last)
#Poisson
#sum(dpois(x=obs[-1],lambda = seir_fit$results$I[-1]*N,log=TRUE))
#Neg Binomial
sum(dnbinom(x=obs[-1],mu=seir_fit$results$I[-1]*N,size=N,log=TRUE))
#SSE
#-sum((obs[-1]-seir_fit$results$I[-1]*N)^2)
}
### logit and its inverse for sigma 
logit <- function (p) log(p/(1-p))    # the logit transform
expit <- function (x) 1/(1+exp(-x))   # inverse logit

f1<-function(par){
f_exp<-expit(par[1])
sigma<-expit(par[2])
parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma =1/duration)
initials <- c(S = (1-(f_exp+I_1/N-R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
-LogLikfun(initials,parameters,dat_csv$totale_attualmente_positivi[c((days-last):days)])
}
est<-optim(fn=f1,par=c(logit(0.0005),logit(0.1)))
#estimated par
expit(est$par)
f_exp<-expit(est$par[1])
sigma<-expit(est$par[2])
parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma = 1/duration)
initials <- c(S = (1-(f_exp+I_1/N-R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
pro<-SEIR(pars = parameters, init = initials, time = 0:last)
#predicted vs observed
cbind(pro$results$I*N,dat_csv$totale_attualmente_positivi[c((days-last):days)])
```

The values of initial parameters are
```{r ini parameters, echo=TRUE}
#mu0
mu0;
#sigma
sigma
#gamma
1/duration
```

```{r status, echo=TRUE}
#mthe unknowrn fraction of exposed people is
f_exp;
#that is 
f_exp/(I_1/N)
#times the infected
```
For the beta parameter, we perfom a simulation on its trend by means of a Bayesian Structural Time Series using the library bsts of R.    
We estimate a BSTS model specifing a local linear trend ans 1000 simulation.

@@

```{r bsts model fit and prediction, echo=TRUE}
# Bayesian Structural Time Series
ss <- AddLocalLinearTrend(list(), beta_vec)
model1 <- bsts(beta_vec,
               state.specification = ss,
               niter = 1000)
par(mfrow = c(1,1))
plot(model1, "components", ylab="Beta1 coefficient", xlab="days")
#previsioni
pred1 <- predict(model1, horizon = 16, burn = 100)
par(mfrow = c(1,1))
plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions")
# matrix of beta coefficients
coef<-(pred1$distribution[,3:16])
par(mfrow = c(1,1))
```
For each vector of simulated beta coefficients we perform a SEIR model.
We save the results and plot a credible interval at 50% (25-75% percentile).

```{r  scenario plot , echo=TRUE}
seir1_sim<-NULL
for(s in 1:dim(coef)[1]){
  # average number of single connections of an infected person
  # less contacts, less probability of new infections
  # we keep constant the other parameters
  forecast<-14
  seir1<-NULL
  for(i in 1:forecast){
    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
    if( i==1) initials <- c(S = 1-(f_exp+I_start/N+R_start/N), E = f_exp, I = I_start/N, R = R_start/N)
    if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
    seir1_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
    seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
  }
  seir1_sim<-rbind(seir1_sim,cbind(rep(s,forecast),seir1))

}
seir1_sim[,2]<-rep(1:forecast,dim(coef)[1])
colnames(seir1_sim)[1]<-"sim"

### confidence limits
I_seir_med<-tapply(seir1_sim$I,seir1_sim$time,median)
I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.25)
I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.75)

days.before<-date[1:days]
days.ahead<-date[(days+1):(days+forecast)]
step.ahead<-forecast+1
mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Italy: Scenario  (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```

At the end of the 2 weeks (`r date[length(date)]`):
*the number of infected is  (`r (tapply(seir1_sim$I,seir1_sim$time,median)*N)[forecast]`).

*the total number of COVID19 cases is expected to be  (`r (tapply(seir1_sim$I,seir1_sim$time,median)*N+tapply(seir1_sim$R,seir1_sim$time,median)*N)[forecast]`).

The estimated (median) numbers of the current scenario by date are:
```{r  scenario numbers, echo=TRUE}
S_seir_med<-tapply(seir1_sim$S,seir1_sim$time,median)
E_seir_med<-tapply(seir1_sim$E,seir1_sim$time,median)
R_seir_med<-tapply(seir1_sim$R,seir1_sim$time,median)
forecast<-data.frame(S_seir_med,E_seir_med,I_seir_med,R_seir_med)*N
colnames(forecast)<-c("Susceptible","Exposed","Infected","Removed")
rownames(forecast)<-days.ahead
kable(forecast)
```
